---
title: "Intro to readxl"
author: "Jen Johansen"
date: "December 7, 2019"
output: html_document
---

# Learning `readxl`

We are working with data in readxl and it's going to be amazing!

We're going to use packages:

- tidyverse
- readxl
- writexl
- here

```{r packages}
library(tidyverse)
library(readxl)
library(writexl)
library(here)

# install.packages ("package_name")
```

### Read in fish counts

```{r}
#### Read in the fish_counts_curated.csv
fish_counts <- read_csv(here("R-workshop-data","fish_counts_curated.csv"))
```

### Read in data from a single Excel worksheet

Now, to read in an excel spreadsheet, we'll use the function 'read_excel()'.

substrate_cover_curated.xlsx

```{r}
substrate_cover <- read_excel(here("R-workshop-data","substrate_cover_curated.xlsx"),
                              skip = 4,
                              na = "-9999")
```

### Use read_excel() to read in a range of values

```{r}
carp_cover_2000 <- read_excel(here("R-workshop-data","substrate_cover_curated.xlsx"), range = "D12:I15",
                              col_name = c("site",
                                           "transect",
                                           "quad",
                                           "side",
                                           "type",
                                           "coverage"))
```

### How can we read in data from an Excel workbook with MULTIPLE worksheets?

We only want to read in data for Goleta Beach (worksheet named: golb)

```{r}
kelp_golb <- read_excel(here("R-workshop-data","kelp_counts_curated.xlsx"),
                        sheet = "golb")
```

### Read in AND combine data from multiple Excel worksheets simultaneously using purrr::map_df()

```{r}
# create a path to get to the data
# kelp_path <- here("R-workshop-data", "kelp_counts_curated.xlsx")
# 
# Comman + Shift + M for new pipe operatttor
# kelp_all_sites <- kelp_path %>%
#   excel_sheets() %>% 
#   set_names() %>% 
#   purr::map_df(read_excel, kelp_path)

```

### Write a file back to an Excel spreadsheet

We're going to write our "kelp_golb" object to a CSV file

```{r}
write_csv(kelp_golb, "kelp_goleta_july.csv")

#write_csv(kelp_golb, here())

write_xlsx(kelp_golb, "kelp_goleta_excel.xlsx")
```

Hello

Hello again
