---
title: "intro-to-ggplot2"
output: html_document
---


## Attach packages

```{r}
library(tidyverse)
library(readxl)
library(here)
```


## Read in the data

```{r}
ca_np <- read_csv(here("R-workshop-data", "ca_np.csv"))

ci_np <- read_xlsx(here("R-workshop-data", "ci_np.xlsx"))
```

## Make a graph

1. Tell R we're using 'ggplot'
2. Tell R what data we're using, and what variables we want to plot
3. Let R and ggplot know what type of grph we're creating (geom)


```{r, eval = FALSE}

ggplot(data = THIS, aes(x = THIS, y = THIS))  +
  geom_THIS()

```

Now let's create a graph of CI visitation:

```{r}
ggplot(data = ci_np, aes(x = year, y = visitors)) +
  geom_line()
```

```{r}
ggplot(data = ci_np, aes(x = year, y = visitors)) +
  geom_point()
```

1. Create a column graph with this data
2. Create an area graph with this data

```{r}
ggplot(data = ci_np, aes(x = year, y = visitors)) +
  geom_col()
```


```{r}
ggplot(data = ci_np, aes(x = year, y = visitors)) +
  geom_area()
```


## Intro to customizing 'ggplot' graphs

- color = (update point or line colors)
- fill = (update fill characteristic of an element)
- linetype =
- pch = (value)
- size = 
- alpha = 


```{r}
ggplot(data = ci_np, aes(x = year, y = visitors)) +
  geom_line(color = "coral4", 
            linetype = "dashed")
```

```{r}
ggplot(data = ci_np, aes(x = year, y = visitors)) +
  geom_point(color = "blue",
             pch = 17,
             size = 4,
             alpha = 0.5)
```

```{r}
ggplot(data = ci_np, aes(x = year, y = visitors)) +
  geom_point(color = "deeppink",
             pch = 24,
             size = 3,
             alpha = 0.5)
```


I have some awesome text here, and below is my amazing finalized graph:

```{r, echo = FALSE}
ggplot(data = ci_np, aes(x = year, y = visitors)) +
  geom_point(color = "deeppink",
             pch = 17,
             size = 3,
             alpha = 0.5)
```

