---
title: "Filter/joins (like VLOOKUP in Excel)"
output: html_document
---


## Attach packages:

tidyverse, readxl, here

```{r, include = FALSE}

library(tidyverse)
library(readxl)
library(here)

```

Read in 3 files and store them as objects

- invert_counts_curated.xlsx
- fish_counts_curated.csv
- kelp_counts_curated.xlsx

```{r}

invert_counts <- read_excel(here("R-workshop-data", "invert_counts_curated.xlsx"))

fish_counts <- read_csv(here("R-workshop-data", "fish_counts_curated.csv"))

kelp_counts_abur <- read_excel(here("R-workshop-data", "kelp_counts_curated.xlsx"))

```

- select() function: keep, exclude or rearrange COLUMNS
- filter() function: keeps or excludes ROWS (observations) based on whether or not they match variable conditions


## Filter to match rows based on a character string

Only keep observations from fish_counts where the common_name variable matches "garibaldi"

```{r}

fish_garibaldi <-fish_counts %>% 
  filter(common_name == "garibaldi")

# Use janitor::clean_names() to automatically update all of your column names to lowercase_snake_case
```


## Filter to keep or exclude observations based on numeric conditions

Create a subset where total_counts is greater than or equal to 50:

```{r}

fish_over50 <- fish_counts %>% 
  filter(total_count >= 50)

```


Create a subset that only contains observations for fish at Mohawk Reef ("mohk"):

```{r}

mohawk_fish <-fish_counts %>% 
  filter(site == "mohk")

```

## Filter to keep rows that match THIS or THAT

Example: We want to keep any observations for garibaldi, blacksmith, or black surfperch

Use the %in% operator to look for any matches within a vector

```{r}

fish_3sp <- fish_counts %>% 
  filter(common_name %in% c("garibaldi", "blacksmith", "black surfperch"))

# Writing the or statements explicitly:
# Use the vertical line | to indicate "or"

fish_3sp_v2 <- fish_counts %>% 
  filter(common_name == "garibaldi" | common_name == "blacksmith" | common_name == "black surfperch")

```

## Filter to return rows that match conditions for MULTIPLE variables (AND statements)

I want to create a subset called urchin_abur_mohk thtat only keeps observations where:

- site is "abur" or "mohk" AND
- species is "purple urchin"

Some ways to do this:

- separate "AND" conditions within a single filter line by a comma
- separate "AND" conditions within a single filter line by an ampersand (&)
- putting each filter condition within a separate filter line

Reminder: "OR" %in%
```{r}
urchin_abur_mohk <- invert_counts %>%
  filter(site %in% c("abur", "mohk"), common_name == "purple urchin")
```





