---
title: "Pivot tables"
output: html_document
---

# Attach libraries

```{r}

library(tidyverse)
library(readxl)
library(here)
library(skimr) # install.packages("skimr")

```

## Read in our data

our data is about lobsters

```{r}
lobsters <- read_xlsx(here("R-workshop-data/lobsters2.xlsx"))
```


Explore our data

```{r}

# head() function show the first 5 lines of data

head(lobsters)

# summary() to give a summary of each column

summary(lobsters)

# skimr::skim() gives additional summary statistics and is pipable (%>%) aka can be used further in analysis

skim(lobsters)

# ggplot of lobsters

ggplot(data = lobsters, aes(x = year))+
  geom_bar() +
  facet_wrap(~site)

```


## Our task:

How has the average size of lobsters has changed over time across sites?

Pivot tables in R are groupy_by %>% summarize


```{r}

lobsters %>% # in my head I say "and then"; i.e. passes the data to the next argument as you continue wrangling it
  group_by(year) %>% 
  summarize(count_by_year = n())

```

What if we didn't group_by first?


```{r}

lobsters %>% 
  summarize(count_by_something = n())

## assign our summary to a new variable

lobsters_sum <- lobsters %>% 
  summarize(count_by_something = n())


```


## group_by multiple variables

And summarize by multiple variables

We used mean() to calculate the mean size of lobsters by each site and year, and used na.rm = TRUE to calculate mean despite having missing values in our raw data.


```{r}

siteyear_summary <- lobsters %>% 
  group_by(site, year) %>% 
  summarize(count_by_siteyear = n(), 
            mean_size_mm = mean(size_mm, na.rm = TRUE),
            sd_size_mm = sd(size_mm, na.rm = TRUE))

siteyear_summary

```


# Activity

1. Calculate the median size_mm (?median)
2. Create and ggsave a plot




```{r}

siteyear_summary <- lobsters %>% 
  group_by(site, year) %>% 
  summarize(count_by_siteyear = n(), 
            mean_size_mm = mean(size_mm, na.rm = TRUE),
            sd_size_mm = sd(size_mm, na.rm = TRUE),
            median_size_mm = median(size_mm, na.rm = TRUE))

siteyear_summary

##ggplot: geom_line

ggplot(data = siteyear_summary, aes(x = year, y = median_size_mm, color = site)) +
  geom_line()


# I created a figures folder fromt he RStudio Files pane in order to be organized and for this code to run
ggsave(here("figures/lobsters-line.png"))

```

# dplyr::count() for counting by group
This is a "shortcut" function

```{r}

lobsters %>% 
  count(site, year)

## this is the same as 

lobsters %>%
  group_by(site, year) %>% 
  summarize(n = n()) %>% 
  ungroup()

```


# mutate() adds a new column (make a new variable)

- add new column that is the size in meters
- add a new column that is a repeated value

```{r}
head(lobsters)

lobsters %>% 
  mutate(size_m = size_mm / 1000,
         millenia = 2000,
         observer = "Allison Horst")

```


## select() to subset variables by columns

let's select
- date
- site
- size_m

```{r}

lobsters_selected <- lobsters %>% 
  mutate(size_m = size_mm / 1000) %>% 
  select(date, site, size_m)

```



